The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-10-01 paper2_JNLS/paper.tex
NM
NNM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC). In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the region of shallower depth, the wave distribution skews markedly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. 
Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.

The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~experimental discoveries guiding theoretical developments and theoretical predictions inspiring new experimental tests. We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied. We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.

%~experimental discoveries guiding model developments, as well as new theoretical predictions inspiring new experimental tests.

%IDEA: Could comment on synergy, or close cooperation, between theory and experiments and the success thereof, then cite Ristroph erosion paper.

% combination
% draw close comparison
% with an even closer comparison between theoretical predictions and experimental measurements.

@@

% ideal, paradigm, canonical, tractable

@@

 % Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Spatial moments}
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the region of shallower depth, the wave distribution skews markedly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. 
Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.

The purpose of the present manuscript is to expand upon the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical, majdaqi2019}, with special emphasis placed on the synergy between experiments and theory. We present a range of previously unpublished experimental data which motivates further refinement of the theory. Likewise, new theoretical predictions have motivated additional experimental measurements, and we present a comparison between the two. This close cooperation between theory and experiments has proven a successful strategy \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}

% New data: slopes, autocorrelation, depth ratio varied.
% We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.

% combination; draw close comparison; with an even closer comparison between theoretical predictions and experimental measurements.

@@

@@

 % Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.9 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Spatial moments}
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
